---
title: "Econometrics Assignment 3"
author: "Chase Bookin & Cole Price"
date: "July 7, 2020"
output:
  pdf_document: default
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(tidycensus)
library(ggthemes)
library(ggplot2)
library(janitor)
library(dplyr)
library(gt)
library(readxl)
library(sf)
library(scales)
library(magrittr)
library(haven)
library(infer)
library(Lahman)
library(xml2)
library(rvest)
library(devtools)
library(broom)
library(retrosheet)
library(skimr)
library(knitr)
library(tinytex)
library(lubridate)
library(pander)
```


```{r, include=FALSE}
capm <- read_xlsx("CAPM.xlsx") %>% 
  mutate(date = ymd(date),
         "Market Risk Premium" = mkt - riskfree,
         "Risk GM" = gm - riskfree,
         "Risk Microsoft" = msft - riskfree,
         "Risk Exxon" = xom - riskfree) %>%
  rename("Date" = date,
         "General Motors" = gm,
         "Microsoft" = msft,
         "Exxon Mobil" = xom,
         "Market Return" = mkt,
         "Risk Free Rate" = riskfree)

capm_summary <- capm %>%
  select(-Date)

panderOptions('round', 5)
skim_with(numeric = sfl(mean, sd))
skim_without_charts(capm_summary)
```

```{r, include=FALSE}
capm_gt <- read_xlsx("CAPM.xlsx") %>% 
  mutate(date = ymd(date)) %>% 
  pivot_longer(
    cols = c(gm, msft, xom),
    names_to = "company",
    values_to = "company_return"
  ) %>% 
  mutate("mkt_risk" = mkt - riskfree,
         "company_risk" = company_return - riskfree)

capm_gt_return <- capm_gt %>% 
  group_by(company) %>% 
  summarize(Mean = mean(company_return),
            SD = sd(company_return),
            Min = min(company_return),
            Max = max(company_return)) %>% 
  mutate(company = ifelse(company == "gm", "General Motors Return",
                          ifelse(company == "msft", "Microsoft Return",
                                 "Exxon-Mobil Return"))) %>% 
  rename(Variable = company)

capm_gt_risk <- capm_gt %>% 
  group_by(company) %>% 
  summarize(Mean = mean(company_risk),
            SD = sd(company_risk),
            Min = min(company_risk),
            Max = max(company_risk)) %>% 
  mutate(company = ifelse(company == "gm", "General Motors Risk Premium",
                          ifelse(company == "msft", "Microsoft Risk Premium",
                                 "Exxon-Mobil Risk Premium"))) %>% 
  rename(Variable = company)

capm_gt_misc <- read_xlsx("CAPM.xlsx") %>% 
  mutate(date = ymd(date),
         "mkt_risk" = mkt - riskfree) %>% 
  pivot_longer(
    cols = c(mkt_risk, riskfree, mkt),
    names_to = "glorp",
    values_to = "misc"
  ) %>% 
  group_by(glorp) %>% 
  summarize(Mean = mean(misc),
            SD = sd(misc),
            Min = min(misc),
            Max = max(misc)) %>% 
  rename(Variable = glorp) %>% 
  mutate(Variable = ifelse(Variable == 'mkt', "Market Return",
                           ifelse(Variable == 'riskfree', "Risk Free Return",
                                  "Market Risk Premium")))

guapo_gt <- bind_rows(capm_gt_return, capm_gt_risk, capm_gt_misc)

  
```

**C) Data Summary: **
```{r, echo=FALSE}
gt(data = guapo_gt) %>% 
  fmt_number(
      columns = vars(Mean, SD, Min, Max),
      decimals = 4
    ) 
```
  
From this summary table, we can see that the average return for General Motors is negative with a value of -0.0122. The other two stocks - Microsoft and Exxon-Mobil - each demonstrate a positive mean return. Additionally, General Motors has the highest standard deviation on its returns, more than double the Exxon-Mobil standard deviation and slightly above Microsoft's standard deviation. We will see that this extra volatility demonstrated by both General Motors and Microsoft as compared to Exxon-Mobil will be mirrored in the securities' beta values. The market return in this period had an average value of 0.0009 with a standard deviation of 0.0464, only slightly lower than Exxon-Mobil's standard deviation. The risk-free return in this period was 0.0026 on average wit ha standard deviation of 0.0014. This extremely low volatility is what we expect from U.S. Treasury Bills.
  
**D) **  
```{r, echo=FALSE}
capm_gt %>%
  filter(company == "msft") %>% 
  ggplot(aes(x = company_risk)) +
  geom_histogram(bins = 50, color = "black", fill = "lightblue") +
  theme_light() +
  scale_x_continuous(breaks = c(-.4, -.3, -.2, -.1, 0, .1, .2, .3, .4)) +
  scale_y_continuous(breaks = c(0, 2, 4, 6, 8, 10, 12, 14)) +
  labs(
    title = "Distribution of Microsoft Risk Premium",
    subtitle = "Using data from 120 monthly observations from January 1999 - December 2008",
    x = "Risk Premium",
    y = "Frequency"
  ) +
  geom_vline(xintercept = -.0003, linetype = "longdash", color = "black") +
  annotate("text", x = -.15, y = 11.5, label = "Average Premium: -0.000294")
```
  
This histogram demonstrates the fairly normal distribution of the Microsoft risk premium. Most of the monthly risk premium data fall in the range from -0.1 to 0.1 with several observations well beyond this mark in both positive and negative directions. The most common range of risk premiums recorded was in the bin just above 0 from about 0 to 0.02.


**E) **
```{r, echo=FALSE}
msft <- capm_gt %>%
  filter(company == "msft")

reg_msft <- lm(company_risk ~ mkt_risk, msft) %>% 
  tidy(conf.int = TRUE)

gm <- capm_gt %>%
  filter(company == "gm")

reg_gm <- lm(company_risk ~ mkt_risk, gm) %>% 
  tidy(conf.int = TRUE)

xom <- capm_gt %>%
  filter(company == "xom")

reg_xom <- lm(company_risk ~ mkt_risk, xom) %>% 
  tidy(conf.int = TRUE)
```
  
Using OLS regression of company risk premium on the market risk premium, we find
that the CAPM beta values for the three companies are as follows: Microsoft:
1.332; General Motors: 1.302; Exxon-Mobil: 0.395. These security beta values
sugest Microsoft and General Motors are more aggressive stocks - with Microsoft
being slightly more aggressive - while Exxon-Mobil presents a more defensive
position.  
  
**F) **  
```{r}
null <- 1
se_msft <- 0.174326
beta_msft <- 1.33228
t_msft <- (beta_msft - null) / (se_msft)
p_msft <- 2*(1 - 0.9716)
# p-value Microsoft 0.0568

se_gm <- 0.22656
beta_gm <- 1.30222
t_gm <- (beta_gm - null) / (se_gm)
p_gm <- 2*(1 - 0.9082)
# p-value GM 0.1836

se_xom <- 0.1005
beta_xom <- 0.3945
t_xom <- (beta_xom - null) / (se_xom)
# p-value Exxon < 0.0001


```

Based on the above calculations, we find the p-value for Microsoft is 0.0568,
which is greater than the alpha value of 0.05 at a 5% significance level.
Therefore, we fail to reject the null hypothesis that the beta for Microsoft is equal to
1 at a 5% significance level. The p-value of General Motors is 0.1836, well
above 0.05; we are unable to reject the null hypothesis that the beta of General
Motors is 1. The p-value of Exxon-Mobil is extremely close to 0 and well below
0.05, meaning we are able to reject the null hypothesis that the beta of
Exxon-Mobil is equal to 1.  

**G) **  

For a company in a given month, a factor that may contribute to the error term is a new product that increases the stock's monthly return above its normal performance. Additionally, a firm's quarterly earnings report may very well affect the stock's returns for a given month, as could a change in leadership at the company.

